\section{Model and Method}
\label{sec:design}

Chiyoda models evacuation as a feedback loop between physical state,
information state, and action. At each step, hazards evolve; agents observe
nearby exits and hazards; beacons, responders, gossip, and intervention
policies update beliefs; agents revise intentions; belief-weighted navigation
selects paths; social-force dynamics move agents; and telemetry records the
macroscopic outcome.

\subsection{Simulation State and Timing}
\label{subsec:state-timing}

The simulation state contains a grid layout, exits, hazards, responders,
beacons, pedestrian agents, spatial indexes, and telemetry collectors. Each run
is seeded and executes a fixed discrete-time horizon. Within a step, Chiyoda
updates hazards before information propagation, then applies interventions
before navigation. This ordering treats communication as a control action that
can change the next route choice rather than as an after-the-fact label on the
trajectory.

\subsection{Agent Decision Loop}
\label{subsec:agent-loop}

Agents are heterogeneous in exit familiarity, information confidence,
susceptibility to gossip, and physiological response. An agent observes local
state, receives messages, updates exit and hazard beliefs, chooses an
intention, plans a route through its believed environment, and moves under a
social-force update. The model therefore separates ground truth from perceived
truth: two agents at the same location can choose different paths because they
carry different exit knowledge, hazard beliefs, and message histories.

\subsection{Belief State}
\label{subsec:beliefs}

Each agent carries a belief vector over exits, hazards, and general danger.
Exit beliefs store existence probability, congestion estimate, freshness,
source credibility, and gossip hop count. Hazard beliefs store position,
severity estimate, radius estimate, freshness, credibility, and hop count.
Agents do not route through the ground truth environment directly; they route
through the subset of exits and hazards represented in their beliefs.

\subsection{Route Choice and Hazard Coupling}
\label{subsec:route-choice}

Route choice is computed from the agent's current belief state. An agent first
selects among exits whose existence probability exceeds the known-exit
threshold, preferring exits with high existence probability, low congestion
estimate, and fresh information. It then plans with A* over the walkable grid.
Each edge has a base traversal cost, a density penalty from the current spatial
index, and a hazard penalty from the agent's hazard beliefs. Ground-truth
hazards still determine exposure, visibility loss, physiological impairment,
and direct observation, but they do not enter route planning unless the agent
has observed or received a corresponding hazard belief. Agents with no known
exit either explore or follow nearby movement depending on rationality and
behavioral state. This separation is the main mechanism that lets the paper
measure information as a causal intervention rather than as an oracle view of
the environment.

\subsection{Communication Channels}
\label{subsec:channels}

The baseline ITED channels are direct observation, beacon broadcast, and
agent-to-agent gossip. Observation is range-limited by visibility and
physiological impairment. Beacons provide high-credibility exit knowledge
within a fixed radius. Gossip is local, probabilistic, and distorted by hop
count and panic state.

Messages are not treated as perfect state synchronization. Each message carries
source credibility, spatial reach, freshness, and content type. Recipients
combine the message with their existing beliefs, which means communication can
increase confidence, correct stale hazard estimates, or reinforce a route that
later becomes congested.

\subsection{Intervention Policies}
\label{subsec:policies}

The intervention policy interface is the experimental control surface. Every
policy decides whether to fire, selects one or more spatial targets, builds a
message, applies it to recipients inside a radius, and emits telemetry. Chiyoda
currently implements seven deterministic baseline policies:

\begin{itemize}
\item \textbf{Static beacon}: periodic local messages from existing signage
or PA beacon locations.
\item \textbf{Global broadcast}: station-wide high-reach communication.
\item \textbf{Responder relay}: mobile high-credibility messages emitted by
released responders.
\item \textbf{Entropy-targeted}: targets agents with highest belief entropy.
\item \textbf{Density-aware}: targets dense local clusters.
\item \textbf{Exposure-aware}: targets agents under high current or cumulative
hazard pressure.
\item \textbf{Bottleneck-avoidance}: targets active bottleneck zones.
\end{itemize}

The repository also includes an optional generated-message policy for bounded
LLM experiments. This policy uses the same intervention hook as the
deterministic policies, but separates target selection from message proposal:
the simulator exposes a bounded state summary, a generator proposes structured
route guidance, and a validator must accept the proposal before it can affect
agent beliefs. This keeps the scientific comparison aligned with the paper's
main framing. A generated message is not treated as helpful because it is
fluent; it is treated as a safety-control action whose downstream exposure,
queueing, belief, and convergence effects must be measured.

Figure~\ref{fig:policy-taxonomy} summarizes the policy space. Figure~\ref{fig:intervention-loop}
shows where the intervention hook sits in the simulation loop.

\begin{figure*}[!t]
\centering
\footnotesize
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.18}
\begin{tabular}{p{0.16\textwidth}p{0.20\textwidth}p{0.20\textwidth}p{0.19\textwidth}p{0.17\textwidth}}
\toprule
Source & Policy & Target signal & Spatial reach & Safety objective \\
\midrule
Static infrastructure & Static beacon & Fixed exit and route priors & Local beacon radius & Improve local belief quality \\
Central authority & Global broadcast & Shared hazard or exit update & Whole population & Rapid common situational awareness \\
Responder authority & Responder relay & High-credibility local report & Responder neighborhood & Correct stale local beliefs \\
Adaptive controller & Entropy-targeted & Highest belief uncertainty & Selected uncertain agents & Reduce uncertainty where it is largest \\
Adaptive controller & Density-aware & Highest local density & Dense clusters & Reduce crowd pressure \\
Adaptive controller & Exposure-aware & Highest hazard load & Exposed agents & Reduce cumulative exposure \\
Adaptive controller & Bottleneck avoidance & Active bottleneck pressure & Queue-adjacent agents & Avoid harmful route convergence \\
\bottomrule
\end{tabular}
\caption{Information intervention policy taxonomy. Policies differ by source,
target signal, spatial reach, and safety objective.}
\label{fig:policy-taxonomy}
\end{figure*}

\begin{figure*}[!t]
\centering
\footnotesize
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.35}
\begin{tabular}{@{}c@{\quad}c@{\quad}c@{\quad}c@{\quad}c@{\quad}c@{\quad}c@{}}
\fbox{\begin{minipage}{0.18\textwidth}\centering Evolve hazards\\and physiology\end{minipage}} &
\(\rightarrow\) &
\fbox{\begin{minipage}{0.18\textwidth}\centering Observe, beacon,\\and gossip\end{minipage}} &
\(\rightarrow\) &
\fbox{\begin{minipage}{0.18\textwidth}\centering Run intervention\\policy\end{minipage}} &
\(\rightarrow\) &
\fbox{\begin{minipage}{0.18\textwidth}\centering Update beliefs\\and telemetry\end{minipage}} \\
&&&&&&\\[-0.4em]
\multicolumn{7}{c}{$\downarrow$}\\[-0.2em]
\multicolumn{7}{c}{%
\fbox{\begin{minipage}{0.80\textwidth}\centering
Update intentions $\rightarrow$ belief-weighted pathfinding $\rightarrow$
social-force movement $\rightarrow$ step telemetry
\end{minipage}}} \\
\end{tabular}
\caption{Intervention hook in the ITED simulation loop. Policies read the
current physical-information state, update recipient beliefs, and emit
before/after telemetry before navigation.}
\label{fig:intervention-loop}
\end{figure*}

\subsection{Information-Safety Metrics}
\label{subsec:metrics}

The core coupled metric is information-safety efficiency:

\[
\eta_{safe} =
\frac{\Delta H^{+} + \Delta A^{+}}
     {1 + P_{exposure} + P_{queue}},
\]

where $\Delta H^{+}$ is the sum of positive entropy reductions over
intervention events, $\Delta A^{+}$ is the sum of positive belief-accuracy
gains, $P_{exposure}$ is recipient-weighted hazard-load pressure, and
$P_{queue}$ is the sum of bottleneck queue pressure observed at intervention
times. Runs with no interventions have ISE zero. ISE is deliberately not a
utility function for all evacuation outcomes. It is a diagnostic for whether a
communication policy purchases useful belief change cheaply in safety terms.

The complementary diagnostic is harmful convergence. HCI rises when entropy
reduction coincides with concentrated exit usage, queueing, and exposure. A
policy can therefore score well on ordinary information outcomes while scoring
poorly on HCI if it makes agents more certain about the same constrained route.
This is the main measurement distinction in the paper: Chiyoda evaluates
communication by its coupled physical-information consequences, not by reach or
entropy reduction alone.

In the current implementation, harmful convergence is computed as

\[
\chi_{harm} =
\frac{B_{exit}(1 + Q_{peak})(1 + \bar{E})}
     {1 + \Delta H_{int}^{+}},
\]

where $B_{exit}$ is the final dominant-exit share, $Q_{peak}$ is the peak
bottleneck queue length, $\bar{E}$ is mean cumulative hazard exposure, and
$\Delta H_{int}^{+}$ is the positive entropy reduction attributable to
interventions. The denominator prevents ordinary information gain from being
mistaken for harm, while the numerator raises the score when a population
converges on a dominant exit under queueing and exposure pressure. Like ISE,
HCI is a comparative study metric rather than a universal safety law.
