\section{Evaluation}
\label{sec:evaluation}

The primary evaluation will run \texttt{study\_information\_control.yaml}, a
multi-seed comparison of no intervention, static beacon broadcast, global
broadcast, responder relay, entropy-targeted communication, density-aware
communication, exposure-aware communication, and bottleneck-avoidance
communication. The generated paper stats currently point at \statStudyName{}
with \statVariantCount{} variants, \statRunCount{} runs, and
\statInterventionEvents{} recorded intervention events.

\subsection{Outcome Measures}
\label{subsec:outcomes}

We evaluate physical outcomes, information outcomes, and coupled tradeoffs.
Physical outcomes include evacuation count, mean and p95 travel time, peak
mean density, peak bottleneck queue, mean hazard exposure, p95 exposure, and
incapacitation count. Information outcomes include initial/final/mean entropy,
entropy reduction, belief accuracy, intervention reach, and accuracy gain.
Coupled outcomes include information-safety efficiency, harmful convergence,
and exit imbalance.

\subsection{Planned Hypotheses}
\label{subsec:hypotheses}

\begin{description}
\item[H1:] Adaptive targeting improves information-safety efficiency compared
with global broadcast because it reduces uncertainty where it is decision
critical rather than synchronizing the full crowd.
\item[H2:] Global broadcast reduces entropy fastest but can increase exit
imbalance and bottleneck pressure.
\item[H3:] Exposure-aware targeting reduces mean hazard exposure more than
entropy-targeted communication in high-severity hazard regimes.
\item[H4:] Responder relay is robust but delayed; its value depends strongly
on release time and responder reach.
\end{description}

\subsection{Current Status}
\label{subsec:current-status}

The code path for intervention studies is implemented and smoke-tested. The
next empirical step is to run the full study, inspect confidence intervals,
and decide which variants need larger seed counts. At PhD-paper quality, the
study should use substantially more than the default five seeds and should
include a sensitivity analysis over hazard severity, population familiarity,
message credibility, and intervention timing.

\todo{Replace this status subsection with full tables, confidence intervals,
effect sizes, and the final information-safety frontier after running the full
study.}
