\section{Implementation}
\label{sec:implementation}

Chiyoda is implemented as a Python research toolkit. Scenario YAML files define
layouts, cohorts, hazards, responders, information parameters, intervention
policies, seeds, and variants. The scenario manager builds a simulation by
loading the layout, constructing agents, hazards, exits, responders,
navigation, spatial indexing, behavior state machines, and optional
intervention policies.

The core implementation lives in \texttt{chiyoda/core/simulation.py}. The
simulation loop records step telemetry, agent telemetry, cell-level maps,
exit flow, bottleneck dynamics, hazard state, gossip events, and intervention
events. Study execution lives in \texttt{chiyoda/studies/runner.py}; it
materializes variants, runs each seed, aggregates tables, and exports study
bundles. Analysis and figure export live in \texttt{chiyoda/analysis/}.

The information intervention layer lives in
\texttt{chiyoda/information/interventions.py}. It is deliberately small:
policies produce target points and messages, while the shared executor applies
messages to recipients and computes before/after entropy and accuracy. This
keeps policy design separate from belief mutation and telemetry.

This organization is part of the research contribution rather than incidental
software structure. Chiyoda is not only a crowd-motion simulator with extra
logging; it exposes emergency communication as a controllable action with
explicit targets, budgets, credibility, validation, and downstream safety
telemetry. That separation is what makes the paper's main comparisons possible:
message quality, recipient choice, route capacity, hazard exposure, and belief
change can be varied independently instead of being collapsed into a single
``informed agent'' assumption.

Generated-message support lives in \texttt{chiyoda/information/llm.py}. It is
implemented as an optional provider-neutral layer rather than as a dependency
of the baseline studies. The layer defines structured message requests,
generated evacuation messages, validation records, a content-addressed JSON
cache, deterministic template generation, replay-only generation, and a live
OpenAI Responses API adapter. The \texttt{llm\_guidance} policy uses
cache-first semantics: if a matching cache record exists, the simulator reuses
the cached message and validation result; otherwise a live provider may
populate the cache in explicitly configured pilot runs. Paper-facing LLM runs
should use replay-only mode after cache population.

The generated-message validator rejects proposals that invent exits, invent
hazards, recommend already congested exits, give conflicting recommend/avoid
instructions, exceed radius or credibility bounds, abstain, omit route
guidance, produce vague text, or report confidence below the configured
threshold. Intervention telemetry records the provider, model, cache key,
cache status, validation status, validation reasons, and generated text so
LLM behavior can be audited separately from downstream evacuation outcomes.

Two implementation details matter for reproducibility. First, A* pathfinding
uses per-step caches for density penalties, edge weights, and identical
start/goal/belief queries. These caches are cleared whenever the simulation
advances to a new navigation phase, so they do not change the cost function;
they only avoid recomputing identical values during the same step. Second,
gossip reports about the same physical hazard are merged into an existing
hazard belief rather than appended as duplicate rumors. This is both a
performance requirement and a modeling requirement: repeated reports should
update credibility, freshness, severity, and radius estimates for the same
hazard, not create an unbounded list of separate hazards.

The paper pipeline mirrors the code pipeline. Study execution emits tables and
figures, \texttt{paper/scripts/gen\_stats.py} converts the exported bundle
into \texttt{stats.tex}, and the LaTeX build injects those values into the
abstract and evaluation. This keeps the manuscript tied to regenerated study
artifacts rather than hand-maintained result counts.
