\section{Introduction}
\label{sec:intro}

Emergency communication is usually treated as a benign input to evacuation:
give people better information and their decisions should improve. That
assumption is plausible at the individual level, but it is not obviously true
at the crowd level. A warning that reduces uncertainty for everyone at once can
also synchronize route choice, push agents toward the same exit, amplify a
bottleneck, or move people through a contaminated region before the physical
hazard has dissipated. The question is therefore not simply whether information
helps, but when, where, and for whom it helps.

This paper studies evacuation as a coupled physical-information system. The
physical side includes layout geometry, social-force movement, bottleneck
dynamics, hazard spread, and physiological impairment. The information side
includes partial exit knowledge, hazard beliefs, signage and PA broadcasts,
responder credibility, local gossip, message decay, and distortion. Chiyoda
connects these layers by making agents plan through their believed world rather
than the ground truth world. They may avoid a hazard they have heard about,
miss an exit they do not know, herd when uncertainty is high, or follow a
stale route because the belief has not yet decayed.

\paragraph{Research question.} We ask when communication improves the
belief-safety tradeoff and when it creates harmful convergence. More
concretely: which intervention policies improve belief accuracy and reduce
uncertainty per unit of induced exposure, queueing, and exit imbalance? The
completed study package compares static signage, global broadcasts, responder
relay, entropy-targeted broadcasts, density-aware targeting, exposure-aware
targeting, and bottleneck-avoidance targeting, then stress-tests timing,
budget, credibility, and repetition.

Our central hypothesis is that emergency communication should be evaluated as a
safety-control action, not as a monotonic information-delivery mechanism.
Messages are therefore judged by their joint effect on belief quality and
downstream movement risk: a policy can reduce uncertainty and still be unsafe
if it concentrates agents at a bottleneck or routes them through hazard
exposure. The same framing covers generated guidance: LLM-produced messages are
evaluated only when they remain bounded, replayable, validator-checked, and
scored with the same ISE and HCI diagnostics as deterministic messages.

\paragraph{Contributions.} This paper makes five contributions.

\begin{enumerate}
\item \textbf{A controllable information-intervention layer.} Chiyoda turns
emergency communication into a first-class policy object rather than a fixed
scenario assumption. Policies can be static, authority-driven, or adaptive to
entropy, density, exposure, and bottleneck pressure.

\item \textbf{A coupled ITED simulation loop.} Agents carry probabilistic
beliefs over exits and hazards, update them through observation, beacon
broadcast, responder relay, and gossip, then route through belief-weighted
costs under social-force dynamics and hazard impairment.

\item \textbf{Information-safety metrics.} We introduce study outputs that
jointly measure entropy reduction, belief accuracy gain, intervention reach,
hazard exposure pressure, queue pressure, exit imbalance, information-safety
efficiency, and harmful convergence.

\item \textbf{A reproducible empirical package.} YAML study definitions
generate multi-seed bundles with Parquet/CSV tables and paper figures. The
empirical evaluation uses a 50-seed primary study and two 30-seed supporting
studies, plus a 900-run robustness grid across hazard severity and population
familiarity. Generated statistics are ingested through \texttt{stats.tex}.

\item \textbf{A bounded generated-guidance extension.} Optional LLM-mediated
guidance is evaluated with the same safety-control metrics, cache/replay
requirements, and validation constraints. The extension tests whether generated
messages improve downstream ISE/HCI, not whether they sound more natural.
\end{enumerate}

Structure-wise, this paper first positions Chiyoda against pedestrian
dynamics, information-aware evacuation, and ABM validation work. It then
describes the model and intervention policies, explains the software pipeline,
presents empirical results and robustness evidence, distinguishes the work
from prior systems, states limitations, outlines future work, and concludes
with the current evidence boundary.
