\section{Limitations}
\label{sec:limitations}

\input{llm_tables}

Chiyoda is currently a stylized research simulator, not an operational
evacuation planner. The station layouts are simplified grids unless imported
from richer geometry. Hazard physics are intentionally lightweight and should
not be read as validated CBRN dispersion models. Physiological impairment uses
piecewise exposure curves that are useful for comparative simulation but not
medical prediction.

The behavioral model is also a hypothesis generator rather than a calibrated
description of a specific crowd. Agents use simplified BDI-style intentions,
local gossip, panic/anxiety states, and familiarity cohorts. These mechanisms
are valuable because they make assumptions inspectable, but the paper must not
claim real-world predictive validity without calibration against evacuation
drills, incident records, controlled VR experiments, or pedestrian trajectory
data.

Finally, entropy is not equivalent to safety. The point of the paper is
precisely that uncertainty reduction can be harmful when it produces synchronized
movement. The proposed information-safety metrics are therefore comparative
diagnostics, not universal welfare functions.

A natural extension is adaptive message generation, including learned or
language-model-mediated guidance. That extension should not be evaluated by
message fluency alone. Under this paper's framing, any generated message is a
safety-control action: it must be replayable, validated against the simulated
state, prevented from inventing unavailable exits or stale hazard claims, and
measured by downstream exposure, queueing, exit balance, and evacuation
outcomes. Without those controls, richer language could make a policy more
persuasive while also amplifying harmful convergence.

\subsection{LLM Extension Protocol}
\label{subsec:llm-extension-protocol}

The repository now includes an optional LLM-guidance policy scaffold, but it is
not part of the completed empirical claims in this paper. Its purpose is to
support a controlled follow-on question: whether generated evacuation messages
add safety value beyond deterministic static, global, entropy-targeted, and
bottleneck-aware baselines. The protocol deliberately treats an LLM as a
bounded message proposer. The simulator exposes a structured state summary,
requires a structured message containing recommended exits, avoided exits,
hazard references, confidence, and optional abstention, and validates the
proposal before it can affect agent beliefs.

This design keeps the extension compatible with the core contribution. A live
API call is not allowed to be the experimental evidence by itself. Paper runs
must use cached or deterministic replay, record the prompt-relevant state and
model metadata, and report rejected messages separately from accepted
interventions. The correct comparison is therefore not whether LLM text sounds
more natural, but whether a validated generated-message policy improves ISE or
reduces HCI under the same information budget and target-selection constraints
as deterministic baselines.

\subsection{Medium LLM Extension Result}
\label{subsec:llm-medium-result}

A first medium-scale LLM extension study applies this protocol to 80 runs:
eight policy variants across ten random seeds. The study includes the
deterministic baselines, a deterministic template generator, three live OpenAI
prompt/validator ablations, and a replay-only verification variant. The live
LLM policies make only eight intervention attempts per run, compared with
32--72 interventions for the deterministic communication policies.

\llmMediumTable

The result is useful but deliberately bounded. The live OpenAI variants
achieve higher information-safety efficiency than the deterministic policies
under this intervention budget, and replay exactly matches the safety-strict
live variant. However, the harmful-convergence index remains high. The
extension therefore supports a cautious paper claim: validated generated
messages can be made replayable and efficient, but richer adaptive language
does not by itself solve crowd synchronization risk.

\subsection{LLM Target-Selection Ablation}
\label{subsec:llm-target-selection}

The medium LLM result also has an important confound: the generated-message
policies use a much smaller recipient budget than the deterministic policies.
To separate generated-message behavior from recipient selection, we ran a
target-selection ablation with deterministic template generation. The provider,
prompt style, validator profile, cadence, radius, and budget are held fixed;
only the target selector changes.

\llmTargetSelectionTable

The ablation shows that target selection is not a minor implementation detail.
Bottleneck, entropy, density, and static target selectors all produce much
higher ISE than the deterministic baselines under a small generated-message
budget, while global and exposure targeting fall close to the deterministic
frontier. Density targeting has the best evacuation count among the generated
variants and the lowest generated-policy HCI, but it still does not beat the
static deterministic beacon on harmful convergence. The implication is that
the LLM extension should not be interpreted as a language-only improvement.
The safety effect is created by a coupled design: bounded messages, validation,
small recipient budgets, and target selection.

\subsection{LLM Regime Robustness Check}
\label{subsec:llm-regime-robustness}

We therefore ran a focused LLM robustness extension over the same
hazard-severity and population-familiarity axes used in the deterministic
regime grid. The extension contains nine regimes, five seeds per regime, one
live OpenAI safety-strict cache-population variant, and one replay-only variant
per regime. Each run uses the same eight-attempt intervention budget as the
medium safety-strict LLM policy.

\llmRegimeTable

The live run produced 360 generated-message events. Safety validation accepted
359 OpenAI messages and rejected one congested-exit recommendation, which was
handled by the deterministic fallback. Replay reproduced the live aggregate
outcomes exactly across evacuated count, ISE, HCI, intervention count, and
recipient count. The robustness check strengthens the reproducibility claim
for generated guidance, but it also sharpens the limitation: HCI rises sharply
with hazard severity. LLM guidance remains safety-efficient under a small
budget, but it still does not solve harmful convergence in severe hazard
regimes.

\llmRegimeComparisonTable

The direct comparison against deterministic robustness policies clarifies the
tradeoff. Under the smaller LLM intervention budget, safety-strict generated
guidance has higher ISE than every deterministic communication policy averaged
across the regime grid. It does not, however, lower HCI relative to the static
beacon, bottleneck-avoidance, or entropy-targeted baselines. This is
consistent with the target-selection ablation: bounded generated guidance can
buy useful belief change efficiently, while the remaining safety problem is
still route convergence under hazard pressure.
