\section{Related Work}
\label{sec:related}

Chiyoda sits between pedestrian dynamics, hazard-coupled evacuation, and
information-aware decision models. Social-force models represent pedestrians
as self-driven particles whose movement emerges from goal attraction,
repulsion, and contact forces~\cite{helbing1995social}. Cellular automata
models instead discretize space and update movement through local transition
rules~\cite{burstedde2001simulation}. Microscopic studies also show how simple
local interaction rules can produce macroscopic crowd outcomes and failure
modes~\cite{moussaid2011simple}. These traditions are useful for evacuation
because physical failure is often a capacity problem: exit flow, bottlenecks,
counterflow, and density-dependent slowdown can dominate individual intent.
Chiyoda uses this physical tradition as the substrate, but the paper's novelty
does not rest on a new crowd-force law.

Hazard-coupled evacuation models add a time-varying environment. Toxic-gas and
gaseous-material simulations show that contaminant spread, perception, and
information transmission can change route choice and evacuation
outcomes~\cite{makmul2020cellular,liu2021toxicgas}. This line of work is
closest to Chiyoda's CBRN-like setting. The difference is experimental focus:
prior models often include information perception as a behavioral factor,
whereas Chiyoda makes emergency communication itself the object of controlled
comparison. Policies are not merely scenario settings; they are interventions
that can be timed, targeted, budgeted, and evaluated.
Guidance models that account for congestion, hazards, and compliance reinforce
the same point: a route recommendation is a control action whose effect depends
on both the physical field and whether agents follow it~\cite{chu2017variable}.

Incomplete-information evacuation models provide the decision-theoretic
counterpart. Bayesian and game-theoretic formulations study route choice under
uncertainty and congestion~\cite{hoogendoorn2004routechoice,wang2023bayesian}, while social-group ABM
surveys emphasize communication, affiliation, and collective behavior as
central modeling concerns~\cite{templeton2023social,senanayake2024agent}.
Chiyoda builds on this direction by making uncertainty observable as telemetry.
Belief entropy, belief accuracy, intervention reach, exposure, queueing, and
exit imbalance are exported together, which lets the paper ask whether a
belief improvement is also a safety improvement.

Open-source pedestrian toolchains already cover much of the physical
simulation and trajectory-analysis stack. JuPedSim and Vadere provide mature
microscopic pedestrian simulation frameworks with published locomotion models
and analysis workflows~\cite{kemlohwagoum2015jupedsim,kleinmeier2019vadere}.
SUMO also exposes pedestrian simulation and JuPedSim coupling in a wider
multimodal traffic environment~\cite{alvarezlopez2018sumo}. PedPy complements
these simulators by analyzing real or simulated trajectory files for density,
velocity, flow, and fundamental-diagram measurements~\cite{schrodter2025pedpy}.
Chiyoda should therefore not be read as a replacement for these systems. Its
role is narrower: it keeps the motion model inspectable while adding a
replayable information-control layer whose outputs can be compared with, or
exported toward, established pedestrian-analysis tooling.

External geometry and hazard data are also better treated as interfaces than
as reinventions. Station layout work can draw from indoor and station mapping
schemas such as OpenStationMap and from GTFS Pathways records for entrances,
levels, and station interiors~\cite{openstationmap2026,gtfs2026pathways}.
Validated fire and smoke studies should use high-fidelity tools such as the
NIST Fire Dynamics Simulator when physical dispersion accuracy is the target
question~\cite{mcgrattan2013fds}. Chiyoda's current hazard layer remains a
stylized comparative field, but its scenario and telemetry formats make it
possible to calibrate or cross-check against richer geometry, pathway, hazard,
and trajectory references without making those external tools part of the core
paper claim.
That separation follows evacuation-model verification guidance: model
components and emergent behaviors should be checked against appropriate
reference data before predictive claims are made~\cite{ronchi2013verification}.

The resulting distinction is methodological. Chiyoda treats communication as
an experimental control surface rather than a binary assumption about whether
agents know enough. It asks how an emergency system should allocate limited
communication budget under uncertainty, and it measures the side effects of
that allocation. This makes message timing, targeting, credibility, and
spatial reach comparable interventions with measurable safety tradeoffs. The
contribution is therefore not a claim that Chiyoda is more realistic than all
prior evacuation models; it is a claim that information interventions can and
should be evaluated as safety-control actions.
