% !TEX program = pdflatex
% Chiyoda -- research paper. arXiv preprint, attributed.
% Build with: make paper

\documentclass[sigconf,nonacm,screen]{acmart}

\title{Chiyoda: Entropy-Guided Information Control for Hazard-Coupled Pedestrian Evacuation}
\subtitle{A Simulation Framework for Measuring When Emergency Communication Helps or Harms}

\author{\href{https://gabrielongzm.com}{Gabriel Ong Zhe Mian}}
\affiliation{%
  \institution{Independent}
  \country{Singapore}
}
\email{gabrielzmong@gmail.com}

\IfFileExists{stats.tex}{\input{stats}}{%
  \newcommand{\statStudyName}{TBD}%
  \newcommand{\statVariantCount}{0}%
  \newcommand{\statRunCount}{0}%
  \newcommand{\statInterventionEvents}{0}%
  \newcommand{\statBestPolicy}{TBD}%
  \newcommand{\statBestEfficiency}{TBD}%
  \newcommand{\statPeakHarmfulConvergence}{TBD}%
}

\begin{abstract}
Emergency evacuation is often treated as a physical routing problem in which
warnings simply provide better information. The harder control problem is that
the same message can improve local beliefs while synchronizing many agents
toward a hazardous or capacity-constrained route. We present \emph{Chiyoda}, an
Information-Theoretic Evacuation Dynamics framework that treats emergency
communication as a controllable safety action in a coupled
physical-information system. Agents carry probabilistic exit and hazard
beliefs, exchange distorted local gossip, receive beacon and responder
messages, suffer physiological impairment under CBRN-like hazards, and route
through belief-weighted paths under social-force crowd dynamics. Chiyoda
evaluates static, global, responder-relay, entropy-targeted, density-aware,
exposure-aware, and bottleneck-avoidance policies with telemetry that links
belief entropy, belief accuracy, exposure, queueing, exit imbalance, and
evacuation time. In a 50-seed primary study with \statRunCount{} runs and
\statInterventionEvents{} intervention events, \statBestPolicy{} has the
highest information-safety efficiency in the tested scenario
(\statBestEfficiency{}), while global and adaptive policies show that broad
reach and entropy reduction are not sufficient safety objectives. A 900-run
robustness grid further shows that static local messaging remains the most
information-safety-efficient tested policy across all hazard and familiarity
regimes, even though the evacuation-count winner varies. A bounded LLM
extension applies the same standard to generated guidance: validated sparse
messages can be highly information-efficient, but equal-budget generated
guidance weakens that advantage and does not eliminate harmful convergence.
The central claim is therefore not that more information or richer language is
always better, but that emergency communication must be evaluated by coupled
belief and safety effects.
\end{abstract}

\keywords{evacuation simulation, pedestrian dynamics, information theory,
agent-based modeling, emergency communication, CBRN, entropy}

\begin{document}
\maketitle

\input{sections/introduction}
\input{sections/claims}
\input{sections/background}
\input{sections/design}
\input{sections/implementation}
\input{sections/evaluation}
\input{sections/robustness_plan}
\input{sections/related}
\input{sections/limitations}
\input{sections/conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}
